---
title: "In-class Exercise 6 - Horizon Plot"
author: "Li Jiayi"
date: "02/24/24"
date-modified: "last-modified"
execute:
  eval: true
  echo: true
  freeze: true
  warning: false
  message: false
code-fold: false
code-summary: "Show the code"
editor: visual
date-format: long
---

## Load Package

Load the [ggHoriPlot](https://rivasiker.github.io/ggHoriPlot/) package

```{r}
pacman::p_load(ggHoriPlot, ggthemes, tidyverse)
```

## Import Data

the csv has already been changed to long format using excel. To convert in R, can use pivot_longer function instead

```{r}
averp <- read_csv("data/AVERP.csv")
```

## Data Preperatin

Date column is imported as a character field, and it should be converted into a date format instead.

```{r}
 averp <- averp %>% mutate(`Date` = dmy(`Date`))
```

another way is to use lubridate

## Plotting

The code chunk below use `geom_horizon` to create a Horizon Plot, `facet_grid` here creates the plot for different Consumer Items

`theme_few()` : use as less number of features possible

`scale_fill_hcl(palette = 'RdBu'` : here creates a diverging color using color palette, only if there is decrease/increase

```{r}
#| fig-width: 12
#| fig-height: 10
#| code-fold: true
#| code-summary: "Show the code"
averp %>% 
  filter(Date >= "2018-01-01") %>%
  ggplot() +
  geom_horizon(aes(x = Date, y=Values), 
               origin = "midpoint", 
               horizonscale = 6)+
  facet_grid(`Consumer Items`~.) +
    theme_few() + 
  scale_fill_hcl(palette = 'RdBu') + 
  theme(panel.spacing.y=unit(0, "lines"), strip.text.y = element_text(
    size = 5, angle = 0, hjust = 0),
    legend.position = 'none',
    axis.text.y = element_blank(),
    axis.text.x = element_text(size=7),
    axis.title.y = element_blank(),
    axis.title.x = element_blank(),
    axis.ticks.y = element_blank(),
    panel.border = element_blank()
    ) +
    scale_x_date(expand=c(0,0), date_breaks = "3 month", date_labels = "%b%y") +
  ggtitle('Average Retail Prices of Selected Consumer Items (Jan 2018 to Dec 2022)')
```
