---
title: "Hands-on Exercise 2"
author: "Li Jiayi"
date: "01/12/24"
date-modified: "`r format(Sys.Date(), '%Y-%m-%d')`" 
date-format: long
---

# Content of This Page

-   theory: a summary of [*R for Visual Analytics - Chap 2*](https://r4va.netlify.app/chap02)

    -   a summary of ggplot2 extensions

    -   a dive into ggplot2 extensions

-   practice: some exploration about the dataset

    -   [Qucik access to Some Plotting Exercise below](#plots)

# 1. Getting started

check & load packages

```{r}
pacman::p_load(ggrepel, patchwork, 
               ggthemes, hrbrthemes,
               tidyverse) 
```

Importing the data

```{r}
exam_data <- read_csv("data/Exam_data.csv")
```

# 2. A summary of 4 libraries

+--------------+-------------------------------------------------------------+--------------------------------------------------------------------------------------------------+--------------------------------------------------------------------------------------------------+--------------------------------------------------+
| library      | ggrepel                                                     | ggthemes                                                                                         | hrbrthemes                                                                                       | patchwork                                        |
+==============+=============================================================+==================================================================================================+==================================================================================================+==================================================+
| Challenges   | annotation, especially with large number of data points     | ggplot2 comes with eight [built-in themes](https://ggplot2.tidyverse.org/reference/ggtheme.html) | ggplot2 comes with eight [built-in themes](https://ggplot2.tidyverse.org/reference/ggtheme.html) |                                                  |
+--------------+-------------------------------------------------------------+--------------------------------------------------------------------------------------------------+--------------------------------------------------------------------------------------------------+--------------------------------------------------+
| Descriptions | provides geoms for ggplot2 to repel overlapping text labels | provides some extra themes, geoms, and scales for ggplot2                                        | provides typography-centric themes and theme components for ggplot2                              | preparing composite figure created using ggplot2 |
+--------------+-------------------------------------------------------------+--------------------------------------------------------------------------------------------------+--------------------------------------------------------------------------------------------------+--------------------------------------------------+
|              | `geom_text_repel()`                                         | `theme_economist()`                                                                              |                                                                                                  |                                                  |
|              |                                                             |                                                                                                  |                                                                                                  |                                                  |
|              | `geom_label_repel`                                          |                                                                                                  |                                                                                                  |                                                  |
+--------------+-------------------------------------------------------------+--------------------------------------------------------------------------------------------------+--------------------------------------------------------------------------------------------------+--------------------------------------------------+
|              | [link](https://ggrepel.slowkow.com/)                        | [link](https://cran.r-project.org/web/packages/ggthemes/)                                        | [link](https://cinc.rud.is/web/packages/hrbrthemes/)                                             |                                                  |
+--------------+-------------------------------------------------------------+--------------------------------------------------------------------------------------------------+--------------------------------------------------------------------------------------------------+--------------------------------------------------+
|              | [details below](#ggrepel)                                   |                                                                                                  |                                                                                                  |                                                  |
+--------------+-------------------------------------------------------------+--------------------------------------------------------------------------------------------------+--------------------------------------------------------------------------------------------------+--------------------------------------------------+

# 3. Diving into 4 libraries

## ggrepel

replace `geom_text()` by [`geom_text_repel()`](https://ggrepel.slowkow.com/reference/geom_text_repel.html) and `geom_label()` by [`geom_label_repel`](https://ggrepel.slowkow.com/reference/geom_text_repel.html).

```{r}
#| code-fold: true
#| code-summary: "Show the code"

# Density plot of ENGLISH scores combined for both genders faceted by class
ggplot(data = exam_data, 
       aes(x = ENGLISH, fill = GENDER)) +
  geom_density(alpha = 0.5, color = "black", linewidth = 0.3) + 
  labs(title = "Distribution of English Scores by Class", x = "English Score") +
  theme_minimal() +
  facet_grid(CLASS ~ .) +
  theme(axis.text.y = element_blank(),
        axis.ticks.y = element_blank(),
        legend.text = element_text(size = 8),  
        legend.title = element_text(size = 8))


```

::: panel-tabset
## R Graphics

```{r}
ggplot(data=exam_data, 
       aes(x= MATHS, 
           y=ENGLISH)) +
  geom_point() +
  geom_smooth(method=lm, 
              size=0.5) +  
  geom_label(aes(label = ID), 
             hjust = .5, 
             vjust = -.5) +
  coord_cartesian(xlim=c(0,100),
                  ylim=c(0,100)) +
  ggtitle("English scores versus Maths scores for Primary 3")
```

## ggplot2

```{r}
ggplot(data=exam_data, 
       aes(x= MATHS, 
           y=ENGLISH)) +
  geom_point() +
  geom_smooth(method=lm, 
              size=0.5) +  
  geom_label_repel(aes(label = ID), 
                   fontface = "bold") +
  coord_cartesian(xlim=c(0,100),
                  ylim=c(0,100)) +
  ggtitle("English scores versus Maths scores for Primary 3")
```
:::

## ggthemes

ggplot2 comes with eight [built-in themes](https://ggplot2.tidyverse.org/reference/ggtheme.html), they are: `theme_gray()`, `theme_bw()`, `theme_classic()`, `theme_dark()`, `theme_light()`, `theme_linedraw()`, `theme_minimal()`, and `theme_void()`.

[**ggthemes**](https://cran.r-project.org/web/packages/ggthemes/) provides [‘ggplot2’ themes](https://yutannihilation.github.io/allYourFigureAreBelongToUs/ggthemes/) that replicate the look of plots by Edward Tufte, Stephen Few, [Fivethirtyeight](https://fivethirtyeight.com/), [The Economist](https://www.economist.com/graphic-detail), ‘Stata’, ‘Excel’, and [The Wall Street Journal](https://www.pinterest.com/wsjgraphics/wsj-graphics/), among others.

Refer to this [link](https://ggplot2.tidyverse.org/reference/#themes) to learn more about ggplot2 `Themes`

::: panel-tabset
## R Graphics

```{r}
ggplot(data=exam_data, 
             aes(x = MATHS)) +
  geom_histogram(bins=20, 
                 boundary = 100,
                 color="grey25", 
                 fill="grey90") +
  theme_gray() +
  ggtitle("Distribution of Maths scores") 
```

## ggplot2

```{r}
ggplot(data=exam_data, 
             aes(x = MATHS)) +
  geom_histogram(bins=20, 
                 boundary = 100,
                 color="grey25", 
                 fill="grey90") +
  ggtitle("Distribution of Maths scores") +
  theme_economist()
```
:::

## hrbthems

[**hrbrthemes**](https://cinc.rud.is/web/packages/hrbrthemes/) package provides a base theme that focuses on typographic elements, including where various labels are placed as well as the fonts that are used.

The second goal centers around productivity for a production workflow. In fact, this “production workflow” is the context for where the elements of hrbrthemes should be used. Consult [this vignette](https://cran.r-project.org/web/packages/hrbrthemes/vignettes/why_hrbrthemes.html) to learn more.

## **pathwork**

There are several ggplot2 extension’s functions support the needs to prepare composite figure by combining several graphs such as [`grid.arrange()`](https://cran.r-project.org/web/packages/gridExtra/vignettes/arrangeGrob.html) of **gridExtra** package and [`plot_grid()`](https://wilkelab.org/cowplot/reference/plot_grid.html) of [**cowplot**](https://wilkelab.org/cowplot/) package. In this section, I am going to shared with you an ggplot2 extension called [**patchwork**](https://patchwork.data-imaginist.com/) which is specially designed for combining separate ggplot2 graphs into a single figure.

Patchwork package has a very simple syntax where we can create layouts super easily. Here’s the general syntax that combines:

-   Two-Column Layout using the Plus Sign +.

-   Parenthesis () to create a subplot group.

-   Two-Row Layout using the Division Sign `/`

## combining more theme

# 4. Some Plotting Exercise {#plots}

improve some of the plotting practice from [Hands-on Exercise 1](https://isss608-24jan-jiayili.netlify.app/hands-on_ex/hands-on_ex01/hands-on_ex01) using the 4 new libraries!

## Stacked Bar Chart of Race Distribution by Gender

```{r}
#| code-fold: true
#| code-summary: "Show the code"


```

## Boxplot of English Scores by Class

```{r}
#| code-fold: true
#| code-summary: "Show the code"


```

## Scatterplot of Math and Science Scores

```{r}
#| code-fold: true
#| code-summary: "Show the code"


```

## Density Plot of English Scores by Class

```{r}
#| code-fold: true
#| code-summary: "Show the code"


```

# 

# 5. reference

-   [Patchwork R package goes nerd viral](https://www.littlemissdata.com/blog/patchwork)

-   [ggrepel](https://ggrepel.slowkow.com/)

-   [ggthemes](https://ggplot2.tidyverse.org/reference/ggtheme.html)

-   [hrbrthemes](https://cinc.rud.is/web/packages/hrbrthemes/)

-   [ggplot tips: Arranging plots](https://albert-rapp.de/post/2021-10-28-extend-plot-variety/)

-   [ggplot2 Theme Elements Demonstration](https://henrywang.nl/ggplot2-theme-elements-demonstration/)

-   [ggplot2 Theme Elements Reference Sheet](https://isabella-b.com/blog/ggplot2-theme-elements-reference/)
