---
title: "In-class Exercise 1a: Now You 👀 It!"
author: "Li Jiayi"
date: "01/13/24"
date-modified: "last-modified"
execute:
  eval: true
  echo: true 
  warning: false
editor: visual
date-format: long
---

## Loading R packages

In this hands-on exercise, 2 R packages will be used. They are:

-   [tidyverse](https://www.tidyverse.org)

-   [haven](https://haven.tidyverse.org)

The code chunk used as follows:

```{r}
pacman::p_load(tidyverse,haven)
```

## Filtering PISA data

::: callout-note
## Note

the code under this section filtered PISA data to include only SG data

the dataset is saved as a rds file called `stu_qqq_SG.rds` and loaded back as a dataframe.

only `stu_qqq_SG.rds` will be imported for efficiency purpose for the rest of exercise.
:::

The code chunk below uses [`read_sas()`](https://haven.tidyverse.org/reference/read_sas.html) of [haven](https://haven.tidyverse.org) to import PISA data into R environment.

```{r}
#| eval: false
stu_qqq <- read_sas("data/cy08msp_stu_qqq.sas7bdat")
```

The code chunk below use filter() of dplyr to filter observation in SG only.

```{r}
#| eval: false
stu_qqq_SG <- stu_qqq %>%
  filter(CNT == "SGP")
```

The code chunk below use write_rds() to save the data into rds format.

```{r}
#| eval: false
write_rds(stu_qqq_SG,
          "data/stu_qqq_SG.rds")
```

## Importing filtered PISA data

The code chunk below uses `read_rds()` to import PISA SG data into R environment.

```{r}
stu_qqq_SG <- read_rds("data/stu_qqq_SG.rds")
```
