---
title: "Take-home Exercise 4"
author: "Li Jiayi"
date: "03/06/24"
date-modified: "last-modified"
execute:
  eval: true
  echo: true 
  warning: false
code-fold: true
code-summary: "Show the code"
date-format: long
---

# Project Component

EDA

CDA

Geospatial Analysis

Modeling

-   Time Series Clustering

-   Time Series Prediction

-   Time Serious Network Anlysis

# Data Preparation

Please refer to <https://isss608-24jan-group1.netlify.app/prototype/data_preparation/data_preparation> for complete data preparation process contributed by all members of the team, and the final dataset is shared among all members.

## Loading R packages

```{r}
#| code-fold: false
pacman::p_load(ggiraph, plotly, 
               patchwork, DT, tidyverse,
               readxl, gifski, gapminder,
               plotly, gganimate,networkD3, 
               ggtext, grid, ggnewscale,shadowtext,
               corrplot, ggstatsplot, tidyverse)
```

## **Importing Data**

```{r}
bmi <- read_csv("data/big_mac_combined.csv")
```

```{r}
ggstatsplot::ggcorrmat(
  data = bmi, 
  cor.vars = 2:11)
```

filter specific year to understand the correlation

2000

```{r}
filtered_data <- bmi %>% 
  filter(year == 2000)
```

```{r}
ggstatsplot::ggcorrmat(
  data = filtered_data, 
  cor.vars = 1:11)
```

2000

```{r}
filtered_data <- bmi %>% 
  filter(year == 2005)
```

```{r}
ggstatsplot::ggcorrmat(
  data = filtered_data, 
  cor.vars = 1:11)
```

to filter on specific countries, look at singapore only

```{r}
filtered_data <- bmi %>% 
  filter(country == 'Singapore')
```

```{r}
ggstatsplot::ggcorrmat(
  data = filtered_data, 
  cor.vars = 1:11)
```

look at china

```{r}
filtered_data <- bmi %>% 
  filter(country == 'China')
```

```{r}
ggstatsplot::ggcorrmat(
  data = filtered_data, 
  cor.vars = 1:11)
```

look at Russia

```{r}
filtered_data <- bmi %>% 
  filter(country == 'Russia')
```

```{r}
ggstatsplot::ggcorrmat(
  data = filtered_data, 
  cor.vars = 1:11)
```

To compare 2 countries stat: for example, China, and Russia

```{r}
filtered_data <- bmi %>% 
  filter(country %in% c('Russia', 'China'))
```

```{r}
grouped_ggcorrmat(
  data = filtered_data,
  cor.vars = 1:11,
  grouping.var = country,
  )
```
